Sparse Function-space Representation \\ of Neural Networks

### 2023-05-15 paper/main.tex

\textbf{Linearization gives rise to a Gaussian process}
Our goal is to build a sparse function-space representation of a neural network.
Gaussian processes are an extension of the multivariate normal that characterizes a distribution over
functions in terms of the first two moments.
They are defined by a mean function $\mu(\cdot)$ and a covariance (aka kernel) function $\kappa(\cdot,\cdot)$.
Thus, we need these two functions based on the neural network to build the Gaussian process posterior.
Recent work \citep{khan2019approximate,maddox2021fast} has shown that linearizing approximations in the weight space
lead to function space equivalent approximations.
As Gaussian distributions remain tractable under linear transformations, a linear function in
terms of parameters can be converted from the weight space to the function space (see Ch.~2.1 in \cite{rasmussen2006gaussian}) as follows:\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx_i) \, \vw \implies \mu(\vx_i) = 0 \quad \kappa(\mathbf{x}_i, \mathbf{x}_i') = \frac{1}{\delta} \phi(\vx_i)^\top \phi(\vx_i).
\end{equation}
% The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$.
A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function
% $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$
with the Generalized Gaussian-Newton approximation (GGN).
Following the work of \citep{khan2019approximate}, we can view this approximation as building an approximate linear model of the neural network as $\hat{f}_\weights(\mathbf{x}_i) = \Jac{\weights_*}{\mathbf{x}_i} \, \weights$, where $\Jac{\weights}{\mathbf{x}_i} \coloneqq \nabla_\weights f(\mathbf{x}_i)^\top \in \R^{C \times P}$ is the Jacobian at the MAP.
Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$.
\todo{do we need this sentence about the Hessian to get Laplace-GGN??}
Therefore, the GGN approximate linear model can be used to convert our weight space model to the function space by,
\begin{equation} 
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  0 \quad \text{and} \quad
  \kappa(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\JacT{\weights^*}{\mathbf{x}'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}). With this kernel function, we can then combine it with the dataset $\dataset$ to form the posterior $q(\vf)$. \citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual $\vy$, but a transformation of $\vy$ which they refer to as pseudo-data.\todo{The reader may wonder what is pseudo-data then A: Is this better?} Similarly, \citet{immerImprovingPredictionsBayesian2021} obtain the same covariance function but a different mean because they rely on a first order approximation of the neural network to form a Bayesian GLM model. Both approaches attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. In our method, we wish to make predictions from the derived GP and thus we avoid such adjustments. We now present the general formulation of the sparse functional representation of our trained neural network based on the non-sparse approximation. \todo{this sentence can be avoided probably, A: its what Aidan calls setting the scene}

\section{\our: Sparse function-space represenation of neural networks}
\paragraph{Linearization in the dual parameters}
As stated in \citet[Lemma~1]{csato2002sparse}, the posterior process mean and covariance functions of a GP prior, combined with a likelihood function and $\dataset$ can be stated as:
\begin{equation}  \label{eq:gp_pred}
 \myexpect_{p(f_i \mid\vy)}[f_i] = \vk_{\vx i}^\top \valpha \quad \text{and} \quad
\mathrm{Var}_{p(f_{i}\mid\vy)}[f_i] = \kappa_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
where $f_i = f_\weights(\vx_i)$, the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{m \times m}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and vector $\vbeta \in \R^{m}$, defined as:
\begin{equation}
\label{eq:dual_param}
\alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)],  \quad
\beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] ,
\end{equation}\todo{ More explination of the dual equation 6 the loss landscape link. Perbutation point linked directly to that here. }
%
The parameters $(\alpha_i, \beta_i)$ are the dual parameters that arise in the dual formulations used in, for example, support vector machines \citep{cortes1995support}, whose origins are found in work by \citet{kimeldorf1971some}. The above relations holds for generic likelihoods, and no approximations are involved as the expectation is under the exact posterior. The question is how do we form the dual variables without access to the exact posterior and different approximate inference techniques solve the question in different ways. Bound techniques such as \citet{jaakkola1997variational} attempt to bound the integral thus making the integral tractable, while Expecation Propogation (EP) as used in \citet{csato2002sparse} iteratively finds dual variables but sets them to log of the expectation of the approximate posterior. Meanwhile variational inference  \cite{khan2017conjugate, adam2021dual} find them using the variational approximation of the posterior to solve the integral.

>>>>

\textbf{Linearization gives rise to a Gaussian process}
Our goal is to build a sparse function-space representation of a neural network.
Gaussian processes $f$ are an extension of the multivariate normal that characterizes a distribution over
functions in terms of the first two moments.
They are defined by a mean function $\mu(\cdot)$ and a covariance (aka kernel) function $\kappa(\cdot,\cdot)$.
Thus, we need these two functions based on the neural network to build the Gaussian process posterior.
Recent work \citep{khan2019approximate,maddox2021fast} has shown that linearizing approximations in the weight space
lead to function space equivalent approximations.
As Gaussian distributions remain tractable under linear transformations, a linear function in
terms of parameters can be converted from the weight space to the function space (see Ch.~2.1 in \cite{rasmussen2006gaussian}) as follows:\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}_{i}) = \phi^\top\!(\vx_i) \, \vw \implies \mu(\vx_i) = 0 \quad \kappa(\mathbf{x}_i, \mathbf{x}_i') = \frac{1}{\delta} \phi(\vx_i)^\top \phi(\vx_i).
\end{equation}
% The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$.
A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN \citep{khan2019approximate, daxberger2021laplace, maddox2021fast}. The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function
% $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i)$
with the Generalized Gaussian-Newton approximation (GGN).
Following the work of \citep{khan2019approximate}, we can view this approximation as building an approximate linear model of the neural network as $\hat{f}_\weights(\mathbf{x}_i) = \Jac{\weights_*}{\mathbf{x}_i} \, \weights$, where $\Jac{\weights}{\mathbf{x}_i} \coloneqq \nabla_\weights f(\mathbf{x}_i)^\top \in \R^{C \times P}$ is the Jacobian at the MAP.
Using the Hessian of the approximate model, we arrive at the Laplace-GGN posterior for $q(\vw)$.
\todo{do we need this sentence about the Hessian to get Laplace-GGN??}
Therefore, the GGN approximate linear model can be used to convert our weight space model to the function space by,
\begin{equation} 
\label{eq-laplace-approx-function-space}
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx_i) =  0 \quad \text{and} \quad
  \kappa(\mathbf{x}_i, \mathbf{x}_i')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}_i}\JacT{\weights^*}{\mathbf{x}_i'},
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}). With this kernel function, we can then combine it with the dataset $\dataset$ to form the posterior $q(\vf)$. \citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual $\vy$, but a transformation of $\vy$ which they refer to as pseudo-data.\todo{The reader may wonder what is pseudo-data then A: Is this better?} Similarly, \citet{immerImprovingPredictionsBayesian2021} obtain the same covariance function but a different mean because they rely on a first order approximation of the neural network to form a Bayesian GLM model. Both approaches attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. In our method, we wish to make predictions from the derived GP and thus we avoid such adjustments. We now present the general formulation of the sparse functional representation of our trained neural network based on the non-sparse approximation.

\section{\our: Sparse function-space represenation of neural networks}
\paragraph{Linearization in the dual parameters}
As stated in \citet[Lemma~1]{csato2002sparse}, the posterior process $f_{\textrm{post}}$ mean and covariance functions of a GP prior, combined with a likelihood function and $\dataset$ can be parameterized:
\begin{equation}  \label{eq:gp_pred}
 \mu(\vx_i)_\textrm{post} = \vk_{\vx i}^\top \valpha \quad \text{and} \quad
  \kappa(\mathbf{x}_i, \mathbf{x}_i')_\textrm{post} = \kappa_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
, the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{m \times m}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $\kappa_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters consist of the vector $\valpha \in \R^{m}$ and vector $\vbeta \in \R^{m}$, defined as:
\begin{equation}
\label{eq:dual_param}
\alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)],  \quad
\beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] ,
\end{equation}\todo{ More explination of the dual equation 6 the loss landscape link. Perbutation point linked directly to that here. }
%
The parameters $(\alpha_i, \beta_i)$ are the dual parameters that arise in the dual formulations used in, for example, support vector machines \citep{cortes1995support}, whose origins are found in work by \citet{kimeldorf1971some}. The above relations holds for generic likelihoods, and no approximations are involved as the expectation is under the exact posterior. The question is how do we form the dual variables without access to the exact posterior and different approximate inference techniques solve the question in different ways. Bound techniques such as \citet{jaakkola1997variational} attempt to bound the integral thus making the integral tractable, while Expecation Propogation (EP) as used in \citet{csato2002sparse} iteratively finds dual variables but sets them to log of the expectation of the approximate posterior. Meanwhile variational inference  \cite{khan2017conjugate, adam2021dual} find them using the variational approximation of the posterior to solve the integral.
